Exploring Design Space for Differentially Private Bayesian Inference

### 2019-01-04 ICML2019/main.tex
another pass

\begin{abstract} Bayesian inference is a statistical method which
allows to derive a \emph{posterior} distribution,
starting from a \emph{prior} distribution and observed data. Several
approaches have been explored in order to make this process differentially
private. For example \cite{dimitrakakis2014robust} and
\cite{wang2015privacy} proved that, under specific conditions, sampling from
the posterior distribution is already differentially private, while \cite{zhang2016differential} and \cite{foulds2016theory}
designed differentially private mechanisms that output a
representation of the full posterior distribution.

When the output of a differentially private mechanism is a probability
distribution, accuracy is naturally measured by means of
\emph{probabilistic distances} measuring how far this distribution is
from the original one. Some classical examples are total variation distance,
Hellinger distance, $\chi^2$-distance, KL-divergence, etc.

In this work, we explore the design space for differentially private bayesian inference, by considering different metrics and different algorithms we design a mechanism. 
We focus on two discrete models, the Beta-Binomial and the
Dirichlet-multinomial models, and one probability distance, Hellinger distance. 
For different algorithms, we explore three main mechanisms based on the Laplace algorithm family and exponential algorithm family.
For different metrics, we apply them by calibrating noise to the sensitivity of that utility function, where the utility function is the probability metric we want to
use to measure accuracy. To show the usefulness of these mechanisms we
show an experimental analysis comparing their performance.
\end{abstract}
 % TODO: replace with your keywords

@@

\section{Problem Setting}

\subsection{Preliminaries}
\label{sec_background}

\noindent \textbf{Bayesian Inference.}

Given a prior belief $\Pr(\theta)$ on some parameter $\theta$,
and an observation $\dataobs$, the posterior distribution on $\theta$ given $\dataobs$ is defined as:
\[
  \Pr(\theta | \dataobs) \equiv \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
\]
where the expression $\Pr(\dataobs | \theta)$ denotes the
\emph{likelihood} of observing $\dataobs$ under a value of
$\theta$. Since we consider $\dataobs$ to be fixed, the likelihood is
a function of $\theta$.
For the same reason $\Pr(\dataobs)$ is a constant independent of $\theta$.
Usually in statistics the prior distribution $\Pr(\theta)$ is chosen so that it represents
the initial belief on $\theta$, that is, when no data has been observed. In practice though,
prior distributions and likelihood functions are usually chosen so that the posterior
belongs to the same \emph{family} of distribution of the prior. In this case we say that the prior
is conjugate to the likelihood function. Use of a conjugate prior
simplifies calculations and allows for inference to be performed in a
recursive fashion over the data.
% Then, we have:
% \[
%   \bysinfer(\Pr(\dataobs | \theta), \Pr(\theta), \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
% \]

\noindent \textbf{Beta-binomial System.}

In this work we will consider a specific instance of Bayesian inference and one of its generalizations.
We will consider the situation where the underlying data is drawn from a binomial distribution
($\thicksim \binomial{n}{\theta}$), where $\theta$ represents
the parameter --informally called \emph{bias}-- of a Bernoulli
distributed random variable and $n$ is the number of draws. The
prior distribution over $\theta\in [0,1]$ is going to be a beta
distribution, $\betad(\alpha, \beta)$, with parameters
$\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
\[
  \Pr(\theta)\equiv \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
\]
where $\betaf(\cdot,\cdot)$ is the beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ binary values, that is $\dataobs= (x_1,\dots x_n), x_i\in\{0,1\}$, and the likelihood function is:
\[
  \Pr(\dataobs | \theta)\equiv \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
\]
where $\Delta \alpha = \displaystyle\sum_{i=1}^{n}x_i$.
The posterior distribution over $\theta$ can be easily proven to be equal to:\[\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)\]
From now on we will denote by $\bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$ the previous
posterior distribution. This notation will prove itself handy when its arguments will change types in the next example and
in section \ref{} when the private mechanisms will be described.

\noindent \textbf{Dirichlet-multinomial Systems.}

@@

\noindent \textbf{Differential Privacy.} 
\begin{definition}
\label{def_epsilon_dp}

A randomized mechanism $\mathcal{M}$ with domain $\mathbb{N}^{|\mathcal{X}|}$ and range $\range$, is $(\epsilon, \delta)$-differential privacy, if for all adjacent
\footnote{Given $\dataobs, \dataobs'\in\{0,1\}^n$  we say that $\dataobs$ and $\dataobs'$ are adjacent and we write, $\adj{\dataobs}{\dataobs'}$, iff\\
$\displaystyle \sum_{i}^{n}\iverson{x_i = x'_i }\leq 1$. } $\dataobs, \dataobs' \in \mathbb{N}^{|\mathcal{X}|}$ and all $\mathcal{S} \subseteq \range$:
\begin{equation*}
\Pr[\mathcal{M}(\dataobs) \in \mathcal{S}] \leq e^\epsilon \Pr[\mathcal{M}(\dataobs') \in \mathcal{S}].
\end{equation*}

@@

\subsection{Technical Problem Statement}
\label{subsec_problem}
We are interested in exploring mechanisms for privately releasing the full posterior
distributions derived in section \ref{sec_background}, as opposed to just sampling from them.
It's worth noticing that the posterior distributions are fully characterized
by their parameters, and the family (Beta, Dirichlet distribution) they belong to. Hence, in case of the
Beta-Binomial model we are interested in releasing a private version of the pair of
parameters $(\alpha',\beta')=(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)$, and
in the case of the Dirichlet-multinomial model we are interested in a private version of
$\valpha'=(\valpha + \Delta \valpha)$. \cite{zhang2016differential} and \cite{xiao2012bayesian}
have already attacked this problem by adding independent Laplacian noise to the
parameters of the posteriors. That is, in the case of the Beta-Binomial system,
the value released would be: $(\tilde\alpha,\tilde\beta)=(\alpha +  \widetilde{\Delta \alpha},\beta + n - \widetilde{\Delta \alpha})$
where $\widetilde{\Delta \alpha}\thicksim \lap{\Delta \alpha}{\frac{2}{\epsilon}}$,
and where $\lap{\mu}{\nu}$ denotes a Laplace random variable with mean $\mu$ and scale $\nu$.
This mechanism is $\epsilon$-differentially private, and the noise is
calibrated w.r.t. to a sensitivity of 2 which is derived by using
$\ell_1$ norm over the pair of parameters. Indeed, considering two
adjacent data observations $\dataobs, \dataobs'$, that, from a unique prior, give rise to two posterior
distributions, characterized by the pairs
$(\alpha',\beta')$ and $(\alpha'',\beta'')$ then
$|\alpha'-\alpha''|+|\beta'-\beta''|\leq 2$.
This argument extends similarly to the Dirichelet-Multinomial system. Details are introduced in Sec. \ref{sec_mechs}.

However, in previous works, the accuracy of the posterior was measured with respect to $\ell_1$ norm. That is, an upper bound was given on
\[
  \Pr[|\alpha - \tilde\alpha| + |\beta - \tilde\beta |\geq \gamma ]
\]
where $(\alpha, \beta), (\tilde\alpha,\tilde\beta)$ are as defined above.
\todo{This
accuracy metric is meaningless when the results released are
distributions rather than numerical values. In contrast, distribution
metrics such as $f$-divergence, Hellinger distance, etc. come into
mind overtly when we are measuring distance between
distributions. }
{\color{blue}im not sure this is a good way to put it...}
This gives us motivation on exploring the design space
of mechanisms by considering different norms (a distribution metric)
to compute the sensitivity and provide guarantees on the accuracy.

>>>>

\begin{abstract}
% Several approaches have been explored in order to make Bayesian inference
% differentially private.
% For instance, \cite{dimitrakakis2014robust} and
% \cite{wang2015privacy} proved that, under specific conditions of the likelihood function and the prior distribution,
% sampling from the posterior distribution, is already differentially private. \cite{zhang2016differential} and \cite{foulds2016theory}
% designed differentially private (d.p.) mechanisms by adding laplacian noise
% to the numeric parameters of the posterior distributions. 
% Accuracy, is usally (\cite{??}) measured by providing upper bounds, holding w.h.p, on the
% $\ell_1$ or $\ell_\infty$-distance between the numeric parameters
% of the released posterior distribution and the actual real ones of posterior distribution.
We explore the design space for differentially private Bayesian
inference.  We focused on two discrete models for parametric Bayesian
inference i.e., the Beta-Binomial and the Dirichlet-multinomial
models. We propose a variation of the exponential mechanism that uses
a scoring function based on a metric over probailistic measures, i.e.,
the Hellinger distance. We compare it with other approaches based on
the perturbation of the parameters of the posterior distribution.  We
measure the accuracy using the Hellinger distance over the posterior
distribution released and the real one as opposed to measuring the
$\ell_1$ or $\ell_{\infty}$ distance on the parameters.  We finally
present an experimental analysis on the accuracy of these mechanisms.
\end{abstract}
 % TODO: replace with your keywords

@@

\section{Preliminaries}
\label{sec_background}
\noindent \textbf{Bayesian inference.} 
Given a prior belief $\Pr(\theta)$ on some parameter $\theta$,
and an observation $\dataobs$, the posterior distribution on $\theta$ given $\dataobs$ is defined as:
\[
  \Pr(\theta | \dataobs) \equiv \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
\]
where the expression $\Pr(\dataobs | \theta)$ denotes the
\emph{likelihood} of observing $\dataobs$ under a value of
$\theta$. Since we consider $\dataobs$ to be fixed, the likelihood is
a function of $\theta$.
For the same reason $\Pr(\dataobs)$ is a constant independent of $\theta$.
Usually in statistics the prior distribution $\Pr(\theta)$ is chosen so that it represents
the initial belief on $\theta$, that is, when no data has been observed. In practice though,
prior distributions and likelihood functions are usually chosen so that the posterior
belongs to the same \emph{family} of distribution of the prior. In this case we say that the prior
is conjugate to the likelihood function. Use of a conjugate prior
simplifies calculations and allows for inference to be performed in a
recursive fashion over the data.
% Then, we have:
% \[
%   \bysinfer(\Pr(\dataobs | \theta), \Pr(\theta), \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
% \]
% \noindent \textbf{Beta-binomial System.}
In this work we will consider a specific instance of Bayesian inference and one of its generalizations.
We will consider the situation where the underlying data is drawn from a binomial distribution
($\thicksim \binomial{n}{\theta}$), where $\theta$ represents
the parameter --informally called \emph{bias}-- of a Bernoulli
distributed random variable and $n$ is the number of draws. The
prior distribution over $\theta\in [0,1]$ is going to be a beta
distribution, $\betad(\alpha, \beta)$, with parameters
$\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
\[
  \Pr(\theta)\equiv \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
\]
where $\betaf(\cdot,\cdot)$ is the beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ binary values, that is $\dataobs= (x_1,\dots x_n), x_i\in\{0,1\}$, and the likelihood function is:
\[
  \Pr(\dataobs | \theta)\equiv \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
\]
where $\Delta \alpha = \displaystyle\sum_{i=1}^{n}x_i$.
The posterior distribution over $\theta$ can be easily proven to be equal to:\[\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)\]
From now on we will denote by $\bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$ the previous
posterior distribution. This notation will prove itself handy when its arguments will change types in the next example and
in section \ref{} when the private mechanisms will be described.
% \noindent \textbf{Dirichlet-multinomial Systems.}

@@

\noindent \textbf{Differential Privacy.}
\emph{Differential privacy} is a strong quantitative notion of statistical privacy proposed by \cite{}.
Differential privacy of a mechanism implies that changing the input of a mechanism slightly only reflects
in a controlled and bounded way on the output distribution. 
Formally, the $\epsilon$ parameter
controls how much two outputs starting from \emph{close} inputs can differ. When two inputs are close we call them \emph{adjacent}.
By adjacency we usually mean a relation over input databases which includes all the pair of databases which differ by at most one row.
The formal definition of $\epsilon$-differential privacy follows.
\begin{definition}
\label{def_epsilon_dp}

A randomized mechanism $\mathcal{M}$ with domain $\mathbb{N}^{|\mathcal{X}|}$ and codomain $\range$, is $\epsilon$-differential private, if for all adjacent
\footnote{Given $\dataobs, \dataobs'\in\{0,1\}^n$  we say that $\dataobs$ and $\dataobs'$ are adjacent and we write, iff
$\displaystyle \sum_{i}^{n}\iverson{x_i = x'_i }\leq 1$. } $\dataobs, \dataobs' \in \mathbb{N}^{|\mathcal{X}|}$ and all $\mathcal{S} \subseteq \range$:
\begin{equation*}
\Pr[\mathcal{M}(\dataobs) \in \mathcal{S}] \leq e^\epsilon \Pr[\mathcal{M}(\dataobs') \in \mathcal{S}].
\end{equation*}

@@

\noindent \textbf{Technical Problem Statement.}
\label{subsec_problem}
We are interested in exploring mechanisms for privately releasing the full posterior
distributions derived in section \ref{sec_background}, as opposed to just sampling from them.
It's worth noticing that the posterior distributions are fully characterized
by their parameters, and the family (Beta, Dirichlet distribution) they belong to. Hence, in case of the
Beta-Binomial model we are interested in releasing a private version of the pair of
parameters $(\alpha',\beta')=(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)$, and
in the case of the Dirichlet-multinomial model we are interested in a private version of
$\valpha'=(\valpha + \Delta \valpha)$. \cite{zhang2016differential} and \cite{xiao2012bayesian}
have already attacked this problem by adding independent Laplacian noise to the
parameters of the posteriors. That is, in the case of the Beta-Binomial system,
the value released would be: $(\tilde\alpha,\tilde\beta)=(\alpha +  \widetilde{\Delta \alpha},\beta + n - \widetilde{\Delta \alpha})$
where $\widetilde{\Delta \alpha}\thicksim \lap{\Delta \alpha}{\frac{2}{\epsilon}}$,
and where $\lap{\mu}{\nu}$ denotes a Laplace random variable with mean $\mu$ and scale $\nu$.
This mechanism is $\epsilon$-differentially private, and the noise is
calibrated w.r.t. to a sensitivity of 2 which is derived by using
$\ell_1$ norm over the pair of parameters. Indeed, considering two
adjacent data observations $\dataobs, \dataobs'$, that, from a unique prior, give rise to two posterior
distributions, characterized by the pairs
$(\alpha',\beta')$ and $(\alpha'',\beta'')$ then
$|\alpha'-\alpha''|+|\beta'-\beta''|\leq 2$.
This argument extends similarly to the Dirichelet-Multinomial system. Details are introduced in Sec. \ref{sec_mechs}.

However, in previous works, the accuracy of the posterior was measured with respect to $\ell_1$ norm. That is, an upper bound was given on
\[
  \Pr[|\alpha - \tilde\alpha| + |\beta - \tilde\beta |\geq \gamma ]
\]
where $(\alpha, \beta), (\tilde\alpha,\tilde\beta)$ are as defined above.
\todo{This
accuracy metric is meaningless when the results released are
distributions rather than numerical values. In contrast, distribution
metrics such as $f$-divergence, Hellinger distance, etc. come into
mind overtly when we are measuring distance between
distributions. }
{\color{blue}im not sure this is a good way to put it...}
This gives us motivation for exploring the design space
of mechanisms by considering different norms. e.g. a norm over space of measures.
to compute the sensitivity and provide guarantees on the accuracy.
